---
title: "p8122_hw1_jsg2145"
author: "Jared Garfinkel"
date: "9/14/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

# Problem 1

Read in data

```{r}
counterfactuals = tibble(individual = c(1:8), 
                         y_0 = c(0, 1, 0, 1, 1, 0, 1, 0),
                         y_1 = c(0, 0, 1, 0, 0, 1, 0, 0))

counterfactuals
```

## Problem a

The treatment group, y_1, can have no causal effect on health outcome if the treatment difference is 0, a beneficial effect if the treatment difference is 1, or a harmful causal effect if the difference is -1, meaning an individual in a counterfactual was cured in the control group, y_0, but sick in the treatment group.

```{r}
treffect = counterfactuals %>% 
  mutate(difference = y_1 - y_0)

treffect
```

## Problem b

The Average Caussal Effect (ACE) is -1/4 = $$E[y_1] - E[y_0]$$

```{r}
aceffect = treffect %>% 
  summarize(sum_y_0 = sum(y_0),
            sum_y_1 = sum(y_1),
            ace = sum_y_1/8 - sum_y_0/8)

aceffect
```

## Problem b

In the treatment assignment, understanding that we cannot know the counterfactual, the association of the treatment with the outcome is -2/4 = $$E[y_1|A = 1] - E[y_0|A = 0]$$

```{r}
trgroup_0 = treffect %>% 
  mutate(tr = c(1, 0, 1, 1, 0, 0, 0, 1)) %>% 
  filter(tr == 0) %>% 
  summarize(sum_y_0 = sum(y_0))

trgroup_1 = treffect %>% 
  mutate(tr = c(1, 0, 1, 1, 0, 0, 0, 1)) %>% 
  filter(tr == 1) %>% 
  summarize(sum_y_1 = sum(y_1))

trgroups = tibble(trgroup_0, trgroup_1) %>% 
  mutate(effect = (sum_y_1 - sum_y_0)/4)

trgroups
```

## Problem d

Create a partition

```{r}
set.seed(719)
group_0 = counterfactuals %>% 
  sample_frac(size = .5, replace = FALSE) %>% 
  mutate("group" = 0)

group_1 = counterfactuals %>% 
  anti_join(group_0) %>% 
  mutate("group" = 1)

groups = bind_rows(group_0, group_1)

groups
```

Find the observed treatment effects.

The observed effect of treatment is -1/4 = $$E[y_1|A = 1] - E[y_0|A = 0]$$

```{r}
treatment_0 = groups %>% 
  filter(group == 0) %>% 
  summarize(sum_y_0 = sum(y_0))

treatment_1 = groups %>% 
  filter(group == 1) %>% 
  summarize(sum_y_1 = sum(y_1))

treatments = tibble(treatment_0, treatment_1) %>% 
  mutate(effect = (sum_y_1 - sum_y_0)/4)

treatments
```

